---
title: "Untitled"
output: html_document
editor_options: 
  chunk_output_type: console
---

Author: Marie Vibeke Vestergaard

This script will generate a disease file for the genealogy analysis:
lbnr coupled to 0 or 1 for each disease of interest

It also assigns family ID to all individuals and generates pairwise kinship coefficients.

Read in packages:
```{r}
library(dplyr)
library(ggplot2)
```

Data pathways:
```{r}
path_registers <- "/ngc/projects2/predict_r/research/data/sds_registers/2024-12"
path_project <- "/ngc/projects2/predict_r/research/projects/0015_Genealogy_across_diseases"
```


## Disease file
Read in all individuals and all A/B diagnoses from both LPR2 and LPR3. Combine:
```{r}
#All individuals -> keep only lbnr and sex
t_person <- read.csv(paste0(path_registers, "/t_person.csv"), sep=";") %>% select(lbnr, C_KON)

#LPR2 diagnoses
t_diag <- read.csv(paste0(path_registers, "/t_diag.csv"), sep=";") %>% 
  filter(!(C_DIAGMOD %in% c("1", "2")) & C_DIAGTYPE %in% c("A", "B")) %>% 
  select(V_RECNUM, C_DIAG)
#Translate V_RECNUM to lbnr
t_adm <- read.csv(paste0(path_registers, "/t_adm.csv"), sep=";") %>% select(K_RECNUM, lbnr) %>% rename(V_RECNUM = K_RECNUM)
t_diag <- left_join(t_diag, t_adm, by = "V_RECNUM") %>% select(lbnr, C_DIAG)
rm(t_adm)

#LPR3 diagnoses
t_lpr3 <- read.csv(paste0(path_registers, "/diagnoser.csv"), sep=";") %>% 
  filter(diagnosetype %in% c("A", "B") & senere_afkraeftet == "Nej") %>% 
  select(dw_ek_kontakt, diagnosekode)
#Translate dw_ek_kontakt to lbnr
t_adm <- read.csv(paste0(path_registers, "/kontakter.csv"), sep=";") %>% select(dw_ek_kontakt, lbnr)
t_lpr3 <- left_join(t_lpr3, t_adm, by = "dw_ek_kontakt") %>% select(lbnr, diagnosekode) %>% rename(C_DIAG = diagnosekode)
rm(t_adm)

#Combine LPR2 and LPR3
t_diag <- rbind(t_diag, t_lpr3); rm(t_lpr3)
```

If a patients has been registered with A/B diagnosis for the concerning disease once, they count as cases.
Both ICD-8 and ICD-10

Add 0/1 columns for each disease for all individuals in register: 
```{r}
#CD
t_CD <- t_diag %>% filter(grepl(paste("DK50", "56301", "56302", "56308", "56309", sep="|"), C_DIAG))
t_CD <- data.frame(lbnr = unique(t_CD$lbnr), CD = rep(1, length(unique(t_CD$lbnr))))
t_person <- left_join(t_person, t_CD, by = "lbnr")
t_person$CD <- ifelse(is.na(t_person$CD), 0, 1) 
rm(t_CD)

#UC
t_UC <- t_diag %>% filter(grepl(paste("DK51", "56319", "56904", sep="|"), C_DIAG))
t_UC <- data.frame(lbnr = unique(t_UC$lbnr), UC = rep(1, length(unique(t_UC$lbnr))))
t_person <- left_join(t_person, t_UC, by = "lbnr")
t_person$UC <- ifelse(is.na(t_person$UC), 0, 1) 
rm(t_UC)

#RA
t_RA <- t_diag %>% filter(grepl(paste("DM05", "DM06", "71219", "71239", "71259", "71299", sep="|"), C_DIAG))
t_RA <- data.frame(lbnr = unique(t_RA$lbnr), RA = rep(1, length(unique(t_RA$lbnr))))
t_person <- left_join(t_person, t_RA, by = "lbnr")
t_person$RA <- ifelse(is.na(t_person$RA), 0, 1) 
rm(t_RA)

#Psoriasis
t_Ps <- t_diag %>% filter(grepl(paste("DL400", "DL401", "DL402", "DL403", "DL404", "DL408", "DL409", "69610", "69619", "69609", sep="|"), C_DIAG))
t_Ps <- data.frame(lbnr = unique(t_Ps$lbnr), Ps = rep(1, length(unique(t_Ps$lbnr))))
t_person <- left_join(t_person, t_Ps, by = "lbnr")
t_person$Ps <- ifelse(is.na(t_person$Ps), 0, 1) 
rm(t_Ps)

#Psoriatic arthritis
t_PsA <- t_diag %>% filter(grepl(paste("DL405", "DM070", "DM071", "DM072", "DM073", "71219", "71239", "71259", "71209", sep="|"), C_DIAG))
t_PsA <- data.frame(lbnr = unique(t_PsA$lbnr), PsA = rep(1, length(unique(t_PsA$lbnr))))
t_person <- left_join(t_person, t_PsA, by = "lbnr")
t_person$PsA <- ifelse(is.na(t_person$PsA), 0, 1) 
rm(t_PsA)

#MS
t_MS <- t_diag %>% filter(grepl(paste("DG35", "34000", "34001", "34008", "34009", sep="|"), C_DIAG))
t_MS <- data.frame(lbnr = unique(t_MS$lbnr), MS = rep(1, length(unique(t_MS$lbnr))))
t_person <- left_join(t_person, t_MS, by = "lbnr")
t_person$MS <- ifelse(is.na(t_person$MS), 0, 1) 
rm(t_MS)

#PSC
t_PSC <- t_diag %>% filter(grepl(paste("DK8301", "57504", sep="|"), C_DIAG))
t_PSC <- data.frame(lbnr = unique(t_PSC$lbnr), PSC = rep(1, length(unique(t_PSC$lbnr))))
t_person <- left_join(t_person, t_PSC, by = "lbnr")
t_person$PSC <- ifelse(is.na(t_person$PSC), 0, 1) 
rm(t_PSC)

#T1D
t_T1D <- t_diag %>% filter(grepl(paste("DE10", "24900", "24901", "24902", "24903", "24904", "24905", "24906", "24907", "24908", "24909", sep="|"), C_DIAG))
t_T1D <- data.frame(lbnr = unique(t_T1D$lbnr), T1D = rep(1, length(unique(t_T1D$lbnr))))
t_person <- left_join(t_person, t_T1D, by = "lbnr")
t_person$T1D <- ifelse(is.na(t_person$T1D), 0, 1) 
rm(t_T1D)

#Asthma
t_Ast <- t_diag %>% filter(grepl(paste("DJ45", "DJ46", "49300", "49301", "49302", "49308", "49309", sep="|"), C_DIAG))
t_Ast <- data.frame(lbnr = unique(t_Ast$lbnr), Ast = rep(1, length(unique(t_Ast$lbnr))))
t_person <- left_join(t_person, t_Ast, by = "lbnr")
t_person$Ast <- ifelse(is.na(t_person$Ast), 0, 1) 
rm(t_Ast)

#Celiac disease
t_CeD <- t_diag %>% filter(grepl(paste("DK900", "26900", sep="|"), C_DIAG))
t_CeD <- data.frame(lbnr = unique(t_CeD$lbnr), CeD = rep(1, length(unique(t_CeD$lbnr))))
t_person <- left_join(t_person, t_CeD, by = "lbnr")
t_person$CeD <- ifelse(is.na(t_person$CeD), 0, 1) 
rm(t_CeD)

#AS
t_AS <- t_diag %>% filter(grepl(paste("DM45", "71249", sep="|"), C_DIAG))
t_AS <- data.frame(lbnr = unique(t_AS$lbnr), AS = rep(1, length(unique(t_AS$lbnr))))
t_person <- left_join(t_person, t_AS, by = "lbnr")
t_person$AS <- ifelse(is.na(t_person$AS), 0, 1) 
rm(t_AS)

#LE
t_LE <- t_diag %>% filter(grepl(paste("DM321", "DM328", "DM329", "73419",  sep="|"), C_DIAG))
t_LE <- data.frame(lbnr = unique(t_LE$lbnr), LE = rep(1, length(unique(t_LE$lbnr))))
t_person <- left_join(t_person, t_LE, by = "lbnr")
t_person$LE <- ifelse(is.na(t_person$LE), 0, 1) 
rm(t_LE)

#Sar
t_Sar <- t_diag %>% filter(grepl(paste("DD86", "13500", sep="|"), C_DIAG))
t_Sar <- data.frame(lbnr = unique(t_Sar$lbnr), Sar = rep(1, length(unique(t_Sar$lbnr))))
t_person <- left_join(t_person, t_Sar, by = "lbnr")
t_person$Sar <- ifelse(is.na(t_person$Sar), 0, 1) 
rm(t_Sar)

#GD
t_GD <- t_diag %>% filter(grepl(paste("DE050", "24200", sep="|"), C_DIAG))
t_GD <- data.frame(lbnr = unique(t_GD$lbnr), GD = rep(1, length(unique(t_GD$lbnr))))
t_person <- left_join(t_person, t_GD, by = "lbnr")
t_person$GD <- ifelse(is.na(t_person$GD), 0, 1) 
rm(t_GD)

#Iri
t_Iri <- t_diag %>% filter(grepl(paste("DH200","DH201","DH208","DH209", "36400", "36401", "36402", "36403", "36404", "36408", "36409", sep="|"), C_DIAG))
t_Iri <- data.frame(lbnr = unique(t_Iri$lbnr), Iri = rep(1, length(unique(t_Iri$lbnr))))
t_person <- left_join(t_person, t_Iri, by = "lbnr")
t_person$Iri <- ifelse(is.na(t_person$Iri), 0, 1) 
rm(t_Iri)

#BD
t_BD <- t_diag %>% filter(grepl(paste("DM352", sep="|"), C_DIAG))
t_BD <- data.frame(lbnr = unique(t_BD$lbnr), BD = rep(1, length(unique(t_BD$lbnr))))
t_person <- left_join(t_person, t_BD, by = "lbnr")
t_person$BD <- ifelse(is.na(t_person$BD), 0, 1) 
rm(t_BD)

#PG
t_PG <- t_diag %>% filter(grepl(paste("DL88", sep="|"), C_DIAG))
t_PG <- data.frame(lbnr = unique(t_PG$lbnr), PG = rep(1, length(unique(t_PG$lbnr))))
t_person <- left_join(t_person, t_PG, by = "lbnr")
t_person$PG <- ifelse(is.na(t_person$PG), 0, 1) 
rm(t_PG)

#TA
t_TA <- t_diag %>% filter(grepl(paste("DM314", sep="|"), C_DIAG))
t_TA <- data.frame(lbnr = unique(t_TA$lbnr), TA = rep(1, length(unique(t_TA$lbnr))))
t_person <- left_join(t_person, t_TA, by = "lbnr")
t_person$TA <- ifelse(is.na(t_person$TA), 0, 1) 
rm(t_TA)

#MG
t_MG <- t_diag %>% filter(grepl(paste("DG700", "DG733", "DG724", sep="|"), C_DIAG))
t_MG <- data.frame(lbnr = unique(t_MG$lbnr), MG = rep(1, length(unique(t_MG$lbnr))))
t_person <- left_join(t_person, t_MG, by = "lbnr")
t_person$MG <- ifelse(is.na(t_person$MG), 0, 1) 
rm(t_MG)

#NO
t_NO <- t_diag %>% filter(grepl(paste("DG36", "DG371", "DG373", "DG374", "DG379", sep="|"), C_DIAG))
t_NO <- data.frame(lbnr = unique(t_NO$lbnr), NO = rep(1, length(unique(t_NO$lbnr))))
t_person <- left_join(t_person, t_NO, by = "lbnr")
t_person$NO <- ifelse(is.na(t_person$NO), 0, 1) 
rm(t_NO)

#SS
t_SS <- t_diag %>% filter(grepl(paste("DM350", sep="|"), C_DIAG))
t_SS <- data.frame(lbnr = unique(t_SS$lbnr), SS = rep(1, length(unique(t_SS$lbnr))))
t_person <- left_join(t_person, t_SS, by = "lbnr")
t_person$SS <- ifelse(is.na(t_person$SS), 0, 1) 
rm(t_SS)

#PBC
t_PBC <- t_diag %>% filter(grepl(paste("DK743", sep="|"), C_DIAG))
t_PBC <- data.frame(lbnr = unique(t_PBC$lbnr), PBC = rep(1, length(unique(t_PBC$lbnr))))
t_person <- left_join(t_person, t_PBC, by = "lbnr")
t_person$PBC <- ifelse(is.na(t_person$PBC), 0, 1) 
rm(t_PBC)

#HH
t_HH <- t_diag %>% filter(grepl(paste("DE063", sep="|"), C_DIAG))
t_HH <- data.frame(lbnr = unique(t_HH$lbnr), HH = rep(1, length(unique(t_HH$lbnr))))
t_person <- left_join(t_person, t_HH, by = "lbnr")
t_person$HH <- ifelse(is.na(t_person$HH), 0, 1) 
rm(t_HH)

```

Save file:
```{r}
write.table(t_person, paste0(path_project, "/Generated_Data/disease_file_IMIDs.csv"), sep=" ", col.names=T, quote=F, row.names=F)
system(paste0("gzip ", paste0(path_project, "/Generated_Data/disease_file_IMIDs.csv")))
```

## Overview of disease data:
```{r}
t_person <- read.csv(paste0(path_project, "/Generated_Data/disease_file_IMIDs.csv.gz"), sep=" ")
head(t_person)

CD <- t_person %>% filter(CD==1)
dim(CD)[1]
dim(CD)[1] / dim(t_person)[1] *100
sum(CD$C_KON=="K") / dim(CD)[1]*100
rm(CD)

UC <- t_person %>% filter(UC==1)
dim(UC)[1]
dim(UC)[1] / dim(t_person)[1]*100
sum(UC$C_KON=="K") / dim(UC)[1]*100
rm(UC)

RA <- t_person %>% filter(RA==1)
dim(RA)[1]
dim(RA)[1] / dim(t_person)[1]*100
sum(RA$C_KON=="K") / dim(RA)[1]*100
rm(RA)

Ps <- t_person %>% filter(Ps==1)
dim(Ps)[1]
dim(Ps)[1] / dim(t_person)[1]*100
sum(Ps$C_KON=="K") / dim(Ps)[1]*100
rm(Ps)

PsA <- t_person %>% filter(PsA==1)
dim(PsA)[1]
dim(PsA)[1] / dim(t_person)[1]*100
sum(PsA$C_KON=="K") / dim(PsA)[1]*100
rm(PsA)

MS <- t_person %>% filter(MS==1)
dim(MS)[1]
dim(MS)[1] / dim(t_person)[1]*100
sum(MS$C_KON=="K") / dim(MS)[1]*100
rm(MS)

PSC <- t_person %>% filter(PSC==1)
dim(PSC)[1]
dim(PSC)[1] / dim(t_person)[1]*100
sum(PSC$C_KON=="K") / dim(PSC)[1]*100
rm(PSC)

T1D <- t_person %>% filter(T1D==1)
dim(T1D)[1]
dim(T1D)[1] / dim(t_person)[1]*100
sum(T1D$C_KON=="K") / dim(T1D)[1]*100
rm(T1D)

Ast <- t_person %>% filter(Ast==1)
dim(Ast)[1]
dim(Ast)[1] / dim(t_person)[1]*100
sum(Ast$C_KON=="K") / dim(Ast)[1]*100
rm(Ast)

CeD <- t_person %>% filter(CeD==1)
dim(CeD)[1]
dim(CeD)[1] / dim(t_person)[1]*100
sum(CeD$C_KON=="K") / dim(CeD)[1]*100
rm(CeD)

AS <- t_person %>% filter(AS==1)
dim(AS)[1]
dim(AS)[1] / dim(t_person)[1]*100
sum(AS$C_KON=="K") / dim(AS)[1]*100
rm(AS)

LE <- t_person %>% filter(LE==1)
dim(LE)[1]
dim(LE)[1] / dim(t_person)[1]*100
sum(LE$C_KON=="K") / dim(LE)[1]*100
rm(LE)

Sar <- t_person %>% filter(Sar==1)
dim(Sar)[1]
dim(Sar)[1] / dim(t_person)[1]*100
sum(Sar$C_KON=="K") / dim(Sar)[1]*100
rm(Sar)

GD <- t_person %>% filter(GD==1)
dim(GD)[1]
dim(GD)[1] / dim(t_person)[1]*100
sum(GD$C_KON=="K") / dim(GD)[1]*100
rm(GD)

Iri <- t_person %>% filter(Iri==1)
dim(Iri)[1]
dim(Iri)[1] / dim(t_person)[1]*100
sum(Iri$C_KON=="K") / dim(Iri)[1]*100
rm(Iri)

BD <- t_person %>% filter(BD==1)
dim(BD)[1]
dim(BD)[1] / dim(t_person)[1]*100
sum(BD$C_KON=="K") / dim(BD)[1]*100
rm(BD)

PG <- t_person %>% filter(PG==1)
dim(PG)[1]
dim(PG)[1] / dim(t_person)[1]*100
sum(PG$C_KON=="K") / dim(PG)[1]*100
rm(PG)

TA <- t_person %>% filter(TA==1)
dim(TA)[1]
dim(TA)[1] / dim(t_person)[1]*100
sum(TA$C_KON=="K") / dim(TA)[1]*100
rm(TA)

MG <- t_person %>% filter(MG==1)
dim(MG)[1]
dim(MG)[1] / dim(t_person)[1]*100
sum(MG$C_KON=="K") / dim(MG)[1]*100
rm(MG)

NO <- t_person %>% filter(NO==1)
dim(NO)[1]
dim(NO)[1] / dim(t_person)[1]*100
sum(NO$C_KON=="K") / dim(NO)[1]*100
rm(NO)

SS <- t_person %>% filter(SS==1)
dim(SS)[1]
dim(SS)[1] / dim(t_person)[1]*100
sum(SS$C_KON=="K") / dim(SS)[1]*100
rm(SS)

PBC <- t_person %>% filter(PBC==1)
dim(PBC)[1]
dim(PBC)[1] / dim(t_person)[1]*100
sum(PBC$C_KON=="K") / dim(PBC)[1]*100
rm(PBC)

HH <- t_person %>% filter(HH==1)
dim(HH)[1]
dim(HH)[1] / dim(t_person)[1]*100
sum(HH$C_KON=="K") / dim(HH)[1]*100
rm(HH)

```

## CD and UC in focus: use our 2-diag definition!
```{r}
path_basefiles <- "/ngc/projects2/predict_r/research/data/basefiles/2025-01" 
t_person <- read.csv(paste0(path_project, "/Generated_Data/disease_file_IMIDs.csv.gz"), sep=" ")
head(t_person)

ibd <- read.csv(paste0(path_basefiles, "/ibd_patients_two_registrations_2025_01_16.csv"), sep=";")
cd <- ibd %>% filter(IBD_reg2=="CD")
uc <- ibd %>% filter(IBD_reg2=="UC")

t_person <- t_person %>% mutate(cd=ifelse(lbnr %in% cd$lbnr, 1, 0), uc=ifelse(lbnr %in% uc$lbnr, 1, 0))
CD <- t_person %>% filter(cd==1)
dim(CD)[1]
dim(CD)[1] / dim(t_person)[1] *100
sum(CD$C_KON=="K") / dim(CD)[1]*100
rm(CD)

UC <- t_person %>% filter(uc==1)
dim(UC)[1]
dim(UC)[1] / dim(t_person)[1]*100
sum(UC$C_KON=="K") / dim(UC)[1]*100
rm(UC)

write.table(t_person, paste0(path_project, "/Generated_Data/disease_file_IMIDs.csv"), sep=" ", col.names=T, quote=F, row.names=F)
system(paste0("gzip ", paste0(path_project, "/Generated_Data/disease_file_IMIDs.csv")))
```

