---
title: "Untitled"
output: html_document
date: "2023-08-04"
editor_options: 
  chunk_output_type: console
---

Author: Marie Vibeke Vestergaard

This script will assign family ID to all individuals and generates pairwise kinship coefficients.

Read in packages:
```{r}
library(dplyr)
library(ggplot2)
library(ggpubr)
```

Data pathways:
```{r}
path_project <- "/ngc/projects2/predict_r/research/projects/0015_Genealogy_across_diseases" 
path_registers <- "/ngc/projects2/predict_r/research/data/sds_registers/2024-12"
```


## Pairs of parents and offspring:
Remove parents of: individuals with only one parental link, two of the same parent, adopted individuals, individuals with same-sex parents and for twins.
```{r}
#Read in data:
t_person <- read.csv(paste0(path_registers, "/t_person.csv"), sep=";")
sum(t_person$mlbnr == ".")
dim(t_person) 
t_temp <-t_person %>% filter(D_FODDATO != ".") 
min(t_temp$D_FODDATO)
max(t_person$D_FODDATO)

#Find twins
twins <- t_person %>% group_by(D_FODDATO, mlbnr, flbnr) %>% 
  summarise(twin = n(), lbnr = paste0(lbnr, collapse= ", ")) %>% 
  filter(mlbnr != "." & flbnr != "." & twin > 1)
table(as.character(twins$twin))

twin_lbnr <- unlist(c(strsplit(twins$lbnr, ", "))) #Remove individual's parental links
t_person <- t_person %>% mutate(mlbnr = ifelse(lbnr %in% twin_lbnr, ".", mlbnr), flbnr = ifelse(lbnr %in% twin_lbnr, ".", flbnr))
sum(t_person$mlbnr == ".") 
rm(twins, twin_lbnr)

#Remove parents that appear as both mother and father:
both <- unique(t_person$mlbnr)[unique(t_person$mlbnr) %in% unique(t_person$flbnr)]
length(both)
t_person$mlbnr <- ifelse(t_person$mlbnr %in% both, ".", t_person$mlbnr)
t_person$flbnr <- ifelse(t_person$flbnr %in% both, ".", t_person$flbnr)

#Find persons with same parents or only with one parental link:
t_person <- t_person %>% mutate(mlbnr = ifelse(mlbnr == flbnr|mlbnr =="." | flbnr == ".", ".", mlbnr), flbnr = ifelse(mlbnr == flbnr|mlbnr =="." | flbnr == ".", ".", flbnr))
sum(t_person$mlbnr == ".") 

#Find persons with same-sex parents
female <- t_person %>% filter(C_KON =="K")
female <- female$lbnr
male <- t_person %>% filter(C_KON =="M")
male <- male$lbnr

t_person <- t_person %>% mutate(remove = ifelse(mlbnr == ".", "already", ifelse((mlbnr %in% male&flbnr %in%male) | (mlbnr %in% female&flbnr %in%female), "yes", "no")))
table(t_person$remove)
t_person <- t_person %>% mutate(mlbnr = ifelse(remove == "yes", ".", mlbnr), flbnr = ifelse(remove == "yes", ".", flbnr)) %>% select(-remove)
sum(t_person$mlbnr == ".") 
rm(female, male)

#Save at this step - will be used to make ped file later (famID, ID, fatherID, motherID, sex)
t_person <- t_person %>% select(lbnr, mlbnr, flbnr, C_KON)
write.table(t_person, paste0(path_project, "/Generated_Data/edge_file_temp_IMID.csv"), sep=" ", col.names=T, quote=F, row.names=F)

#Make edge list with IID and parent ID (one for each parent - excluding individuals without parental links)
t_person <- t_person %>% select(-C_KON)
first <- t_person %>% filter(mlbnr !=".") %>% select(-mlbnr) %>% rename(ID = lbnr, parentID = flbnr)
second <- t_person %>% filter(mlbnr !=".") %>% select(-flbnr) %>% rename(ID = lbnr, parentID = mlbnr)
edge_list <- rbind(first, second); rm(first, second)

#Save
write.table(edge_list, paste0(path_project, "/Generated_Data/edge_file_IMID.csv"), sep=" ", col.names=T, quote=F, row.names=F)

#How many singletons?
sum(!(t_person$lbnr %in% c(edge_list$ID, edge_list$parentID)))
2618003/dim(t_person)[1] #0.2662776 -> 26.6%

#How many edges?
dim(edge_list)[1]

#How many individuals included in edges?
length(unique(c(edge_list$ID, edge_list$parentID))) 
7213857/dim(t_person)[1]
```

## Assign family ID to individuals
Using the python package networkx

Move to terminal -> to folder (/ngc/projects2/predict_r/research/projects/0015_Genealogy_across_diseases/Scripts)

Run iqsub

Load anaconda:
module load anaconda3/2021.11

Run the python script components.py (python components.py)


## Make ped file:
fam ID, ID, father ID, mother ID, sex
```{r}
# Read in t_person from before
t_person <- read.csv(paste0(path_project, "/Generated_Data/edge_file_temp_IMID.csv"), sep=" ")

# Read in generated fam IDs
famid <- read.csv(paste0(path_project, "/Generated_Data/ids_IMID.csv"), sep=" ")
famid <- famid %>% filter(famid !=1) #The edge list should not have had header - became the first family

#Family overview:
fam_sizes <- famid %>% group_by(famid) %>% summarise(size = n())

# Merge and change ordering of columns
famid <- famid %>% rename(lbnr = indid)
t_person$lbnr <- as.character(t_person$lbnr)
t_person <- left_join(t_person, famid, by="lbnr") %>% select(famid, lbnr, flbnr, mlbnr, C_KON)

#If singleton, remove that row
sum(is.na(t_person$famid))
t_person <- t_person %>% filter(!is.na(famid))

#Recode 1=male, 2=female
t_person$C_KON <- ifelse(t_person$C_KON=="M", 1, ifelse(t_person$C_KON=="K", 2, "."))

#Save ped file:
write.table(t_person, paste0(path_project, "/Generated_Data/ids_file_IMID.ped"), sep=" ", col.names=T, quote=F, row.names=F)
```

## Look into distribution of families - make plot
Pie: percentage with no relatives, trios/quartets, other and the one big family
Number of generations, number of founders
```{r}
#Look into the largest family
large <- t_person %>% filter(famid == 2)
sum(large$mlbnr == ".") 

#Histogram:
fam_sizes <- fam_sizes %>% mutate(type = ifelse(size == dim(large)[1]+1, "Largest family", ifelse(size <=4, "Trios/quartets", "Other families")))
x_axis <- c("3", rep("", 6), "10", rep("", 9), "20", rep("", 9), "30", rep("", 9), "40", rep("", 9), "50", rep("", 9), "60", rep("", 9), "70", rep("", 5), "80", rep("", 5), "93", rep("", 3), "103", rep("", 4), "5991233")

partb <- ggplot(fam_sizes)+
  geom_bar(aes(x=reorder(as.character(size), size), fill=type))+
  theme_classic()+xlab("Family size")+ylab("Count")+
  theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=1), legend.title = element_blank(), legend.position = c(0.8, 0.8),
        text = element_text(family="sans", size=8))+
  scale_y_log10()+scale_x_discrete(labels = x_axis)

#Pie chart
trio_quartet <- fam_sizes %>% filter(size <=4)
other <- fam_sizes %>% filter(size >4 & size !=5991233)
dim(fam_sizes %>% filter(size ==3)) 
dim(fam_sizes %>% filter(size ==4))

part <- data.frame(Individuals = c(2617999, sum(trio_quartet$size), sum(other$size), 5991233), label = c("Singletons", "Trios/quartets", "Other families", "Largest family"))
part <- part %>% mutate(percent = Individuals / sum(Individuals)*100)
part <- part %>% mutate(percent = paste0(substr(percent, 1, 4), "%"))

parta <- ggplot(part, aes(x="", y=Individuals, fill=label))+
  geom_bar(stat="identity")+
  coord_polar(theta = "y")+
  theme_void()+
  theme(legend.title = element_blank(), text = element_text(family="sans", size=8))+
  geom_text(aes(label = percent), position = position_stack(vjust=0.5), size = 3)

legend <- cowplot::get_legend(parta)
parta2 <- ggarrange(parta+theme(legend.position = "none"), legend, nrow=2, ncol=1)
both_fig <- ggarrange(parta2, partb, labels=c("A)", "B)"), font.label = list(size=10, family="sans"))

ggsave(both_fig, filename = paste0(path_project, "/Illustrations/family_overview_IMID.tiff"), width = 174, height = 120, units="mm", dpi=300)
```

## Calculate kinship coefficient
Using the python package pyAGH

Prepare data for the python package:
```{r}
t_person <- read.csv(paste0(path_project, "/Generated_Data/ids_file_IMID.ped"), sep=" ")

#Exclude links to missing parents (founders): -> both the father and the mother will be ".". pyAGH will generate parental links to "0". 
t_person <- t_person %>% filter(mlbnr != ".")

#No need for sex information:
t_person <- t_person %>% select(-C_KON)

#Change header:
colnames(t_person) <- c("famID", "ID", "DAM", "SIRE")
t_person <- t_person %>% select(famID, ID, SIRE, DAM)

length(unique(t_person$famID)) #247295 families

#Save:
write.table(t_person, paste0(path_project, "/Generated_Data/ids_file_preped_IMID.csv"), sep=",", col.names=T, quote=F, row.names=F)

```


Move to terminal -> to folder (/ngc/projects2/predict_r/research/projects/0015_Genealogy_across_diseases/Scripts)

Run iqsub

Load anaconda:
    module load anaconda3/2021.11

Run the python script relatedness.py (python relatedness_biggest.py and python relatedness_rest.py)

Result is table with columns for all pairwise connections: lbnr1, lbnr2, relatedness coefficient

```{r}
#less dataframe
res1 <- read.csv(paste0(path_project, "/Generated_Data/relatedness_coeffs_less_IMID.txt.qz"), sep=" ", header=F)
head(res1)
write.table(res1, paste0(path_project, "/Generated_Data/relatedness_coeffs_less_IMID.txt"), sep=" ", col.names = F, quote = F, row.names = F)
```

Run the python script remove_duplicates.py (python remove_duplicates.py)

Combine the two data frames:
```{r}
res1 <- read.csv(paste0(path_project, "/Generated_Data/relatedness_coeffs_biggest_no_duplicates_IMID.csv.gz"), sep=" ", header=T)
dim(res1)
res2 <- read.csv(paste0(path_project, "/Generated_Data/relatedness_coeffs_less_no_duplicates_IMID.csv.gz"), sep=" ", header=T)

res <- rbind(res1, res2)
dim(res) #49878224 
rm(res1, res2)

write.table(res, paste0(path_project, "/Generated_Data/relatedness_coeffs_all_no_duplicates_IMID.csv"), sep=" ", col.names = T, quote = F, row.names = F)

# Compress file
system(paste0("gzip ", paste0(path_project, "/Generated_Data/relatedness_coeffs_all_no_duplicates_IMID.csv")))
```

Run the python script relatedness_annotation.py (python relatedness_annotation.py)

Look at annotation:
```{r}
res <- read.csv(paste0(path_project, "/Generated_Data/relatedness_annotation_IMID.txt.gz"), sep=" ", header=T)

dim(res)[1] #54652666

#Remove n_paths >2 -> inbreeding
table(res$n_paths, useNA = "always") 
dim(res %>% filter(n_paths >2))[1]/dim(res)[1]*100 #0.05055014% of pairs have >2 paths
res <- res %>% filter(n_paths <3) 
dim(res)[1] #54625039

#Remove parent_rel >2 -> inbreeding
table(res$parent_rel, useNA = "always")
sum(res$parent_rel >2)/dim(res)[1]*100 #0.0008494273% of pairs removed
res <- res %>% filter(parent_rel <3)
dim(res)[1] #54624575 
table(res$annot, useNA = "always") #Still 4883 with annot = NA. Different type of relationship than the included ones, = 0.0089392%

#Remove annot = NA
res <- res %>% filter(!is.na(annot))
dim(res)[1] #54619692

#Save:
write.table(res, paste0(path_project, "/Generated_Data/relatedness_annotation_clean_IMID.csv"), sep=" ", col.names = T, quote = F, row.names = F)
system(paste0("gzip ", paste0(path_project, "/Generated_Data/relatedness_annotation_clean_IMID.csv")))

```
